---
title: "Lab 3 Report"
author: "MaryJo Nelson"
date: "`r format(Sys.Date())`"
output: github_document
---

## Introduction

Researchers studying the number of electric fish species living in various parts of the Amazon basin were interested in whether the presence of tributaries affected the local number of electric fish species in the main rivers (Fernandes et al. 2004).

They counted the number of electric fish species above and below the entrance point of a major tributary at 12 different river locations.

The data is provided in your GitHub repository.

For each question below, write a sentence answering the question and show the code you used to come up with the answer, if applicable.

```{r}

library(tidyverse)
fish <- read_csv("chap12q19ElectricFish.csv")

fish_long <- 
  pivot_longer(fish, speciesUpstream:speciesDownstream,
               names_to = "location",
               values_to = "species") %>% 
  mutate(location = str_remove(location, c("species"))) %>% 
  print()

```

## Question A

> What is the mean different in the number of species between areas upstream and downstream of a tributary? What is the 95% confidence interval of this mean difference. Show your code and write a sentence giving your answer

ANSWER

-   The mean difference is 1.83334
-   The 95% confidence interval of the mean difference is -4.587031 \< x \< 8.253697
-   I got these results by performing a t-test

```{r}

ttest <- t.test(formula = species ~ location, data = fish_long)
ttest

diff(ttest$estimate)

```

## Question B

> Test the hypothesis that the tributaries have no effect on the number of species of electric fish.

ANSWER

-   Based on the t-test, we can fail to reject the hypothesis that tributaries have no effect on the number of species of electric fish for the following reasons:

    -   The p-value is greater than the assumed alpha of 0.05

    -   The confidence interval encompasses 0, meaning that the difference in the means could be 0

-   We can also conclude that the tributaries don't seem to have an effect on the number of species present based on a graph showing the means and confidence intervals of both locations.

    -   The graph shows that there is significant overlap between the confidence intervals for areas upstream and downstream of tributaries, indicating that there is no real difference in the means.

```{r}

fish_summary <-
  fish_long %>% 
  group_by(location) %>% 
  summarize(
    n = n(),
    mean = mean(species),
    sd = sd(species),
    sem = sd/sqrt(n),
    upper = mean + 1.96 * sem,
    lower = mean - 1.96 * sem
  ) 


fish_long %>% 
  ggplot(aes(x = location, y = species)) +
  geom_jitter(aes(color = location), 
              shape = 16, size = 3, 
              alpha = 0.3, width = 0.4) +
  geom_errorbar(aes(y = mean, ymax = upper, ymin = lower), 
                data = fish_summary, 
                width = .1, size = .8) +
  geom_point(aes(y = mean), 
             data = fish_summary, 
             size = 3) +
  scale_color_manual(values = c("darkorange","cyan4")) +
  theme_minimal() +
  guides(color = "none")

```

## Question C

> State the assumptions that you had to make to complete parts (A) and (B). Create a graph to assess whether one of those assumptions was met.

ANSWER

-   The assumptions we had to make were that the sampling units were randomly sampled from the population and that the paired differences have a normal distribution in populations.

-   The distribution appears not to be normal, but rather is skewed right, meaning that that assumption was not met.

```{r}

fish_long %>% 
  ggplot(aes(x = species)) +
  geom_histogram(
    aes(fill = location), 
    bins = 8, 
    alpha = 0.5, 
    position = "identity"
  ) +
  scale_fill_manual(values = c("darkorange","cyan4")) +
  theme_minimal()

```
